### 2022-06-24 manuscript.tex
Merge pull request #244 from SORTEE-Github-Hackathon/emmajhudgins-patch-1 [ci skip]
This build is based on
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/b78a03a54df81c17eb98b9ca80e842dcb1883ab1.

This commit was created by the following CI build and job:
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/b78a03a54df81c17eb98b9ca80e842dcb1883ab1/checks
https://github.com/SORTEE-Github-Hackathon/manuscript/actions/runs/2552679395

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/28647d1b51aa7df9b1214fa3fc7d0743b00d665e/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/28647d1b51aa7df9b1214fa3fc7d0743b00d665e}{SORTEE-Github-Hackathon/manuscript@28647d1}
on June 23, 2022.

@@

Most scientists, including ecologists and evolutionary biologists, are increasingly dependent on computational tools in their research\textsuperscript{\protect\hyperlink{ref-fJWFe93e}{1}}.
Researchers write and use software packages or data analysis code (hereafter, code) as part of their scientific workflow to perform a wide-variety of tasks ranging from data management, data analysis, and study replication, to the application and the development of tools for hypothesis testing.
To maintain code for scientific collaboration requires an efficient and well-documented work-flow\textsuperscript{\protect\hyperlink{ref-1Kqna6l2}{2}}.
To facilitate this process, scientists have been increasingly adopting tools from information and systems technology, such as cloud-based services for documentation and version control (\emph{e.g.}, from the Google Suite, with Docs, Sheets, and Drive; the Microsoft Suite, with Word, Excel, and OneDrive; and GitHub;\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3}}).
However, most researchers lack exposure to adequate code development practices and are required to dedicate valuable time and effort to self-teach the use of research-facilitating tools, and thus may be limited in their ability to adhere to adequate standards of scientific code quality and maintenance\textsuperscript{\protect\hyperlink{ref-fJWFe93e}{1},\protect\hyperlink{ref-10SpoByIw}{4},\protect\hyperlink{ref-O6UbstGG}{5}}.
Here, we review and discuss one of the most used web-based platforms for computational version control and collaboration, GitHub, and provide researchers in ecology and evolutionary biology (EEB) with practical workflows aimed at facilitating their scientific code and management process.

With over 73 million registered users, as of 2022, GitHub is the most widely-used web platform for collaborating on computer code\textsuperscript{\protect\hyperlink{ref-nwCtHDCn}{6}}.
GitHub provides a simplified but powerful web interface that allows users to participate in projects, contribute code, report and discuss code bugs, discover existing code and data, and publish new code.
Through version control, users have a detailed, chronological record of the files and directories stored in their repositories (see \protect\hyperlink{definitions}{Box 1};\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}).
This workflow provides a strong and clear advantage over receiving, processing and sending files back-and-forth (e.g.~via email), a process that can easily become challenging and time-consuming in projects extending in time and in the number of collaborators\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8}}.
Through the combination of version control management and the network- and collaboration-based features, GitHub can broadly facilitate openly available source code alongside concomitant collaborative development\textsuperscript{\protect\hyperlink{ref-kEX5dgzK}{9}}.

Git is the version control system that enables all the collaborative tools available on GitHub.
Although the understanding of basic concepts of Git (such as commit, push, pull, checkout; see \protect\hyperlink{definitions}{Box 1}) is necessary, the GitHub web-based platform and its integrated development environments (such as the GitHub Desktop) allow users to perform most repository and data management operations without opening Git command-line sessions.
Still, GitHub users are encouraged to explore and improve their proficiency in Git when feeling comfortable, so they can feel confident in performing more flexible, complex operations (such as integrating changes from one branch into another branch with \texttt{git\ merge} and \texttt{git\ rebase}). Examples of extensive explanations on Git can be found in journal articles\textsuperscript{\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-PlcxShQU}{10}}, video tutorials\textsuperscript{\protect\hyperlink{ref-NIS0JOW0}{11}}, and books\textsuperscript{\protect\hyperlink{ref-13jOlVcpp}{12}}.

The expansive GitHub user-community and the numerous resources on how to use GitHub have boosted its growing popularity\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-13jOlVcpp}{12}/,\protect\hyperlink{ref-u5aEVE4B}{13}}.
Nevertheless, although multiple articles have encouraged researchers in EEB to adopt GitHub as part of their research process\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-3DKwn1sY}{14}}, its use is still not widespread.
First-time users without formal training in information technology may face steep learning curves because GitHub and its features have been centered on collaboration for code development in information systems\textsuperscript{\protect\hyperlink{ref-139b0pSGc}{15}}.
Moreover, domain-specific perspectives and resources providing tractable examples and practical guidance for researchers in EEB on GitHub are scarce (but see https://ourcodingclub.github.io; https://www.openscapes.org).
A common adoption of GitHub for collaborating on a variety of research tasks can ultimately enable EEB researchers to spend less time on creating novel processes for collaboration and more time on their research\textsuperscript{\protect\hyperlink{ref-ydrk01SR}{16}}.
More importantly, increasing the availability of data and code management standards -- of which GitHub is one increasingly important component -- make research more reproducible and collaborative\textsuperscript{\protect\hyperlink{ref-13QX8XU3J}{17}}.

This paper is the result of an academic hackathon held during the 2021 conference for the Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology (SORTEE, https://www.sortee.org).
We convened a group of around thirty researchers in EEB with varying levels of familiarity with using GitHub as part of their research projects to showcase and discuss how existing features can contribute to documentation and collaboration in EEB research.
During the hackathon, we identified the need for a formal discussion on how EEB researchers can benefit from GitHub and its features to make research more collaborative and transparent.
Here, we outline twelve practical ways that EEB researchers can use GitHub features for more collaborative, transparent, and reproducible science.
We also provide critical perspectives on features that could be improved and catered towards research development.

@@

\textbf{fork}: A fork is a copy of a \textbf{repository} hosted on GitHub. If a repository is public, then anyone can make a fork. Even if they do not have access to \textbf{push} to the original repository, they can make a fork and edit it independently. Forks are linked to the original GitHub repository and ``upstream'' changes (i.e.~those in the original repository) can be \textbf{merged} to keep the fork up to date with the original project. Changes made in the fork can be integrated into the original project via \textbf{pull requests}. \\
\textbf{clone}: Cloning a \textbf{repository} is a way of making a local copy (i.e.~on your computer) of a GitHub \textbf{repository}. If you have access to \textbf{push} to a \textbf{repository}, this can be a first step to contributing to a project. \\

@@

\textbf{commit}: Commits are snapshots of the development of a project. In Git, versions of files and directories are uniquely identified as ``commits'', allowing one to identify and track modifications line-by-line. Commits can include changes in multiple files and must include a brief commit message describing the changes made. A typical workflow is to make some related changes in files, add a commit message (e.g.~``generate and include fig1 in results''), and after several commits \textbf{push} those commits to the remote (\emph{i.e.}, cloud-based) GitHub \textbf{repository}. \\

@@

An EEB research compendium includes all computational materials related to research production, including data, code for analyses and protocols.
Having copies of these files safely stored is essential to protect against accidental modifications or deletions.
Many researchers begin using GitHub as a means to store (or backup) their research compendium\textsuperscript{\protect\hyperlink{ref-MwwMapRG}{18}} to a centralized, readily-available remote server (see \protect\hyperlink{definitions}{Box 1}).
A centralized research compendium stored in a version-controlled repository has the advantages of facilitating collaboration, integrating data and code archiving services, allowing file versions to be accessed and restored, and contributes to open science (see sections below).

GitHub limits committed file sizes to 100 Mb (megabytes) (\textsuperscript{\protect\hyperlink{ref-1Co6ZZjF1}{19}}), which can make it challenging for centralizing research compedia containing larger file sizes.
Users may still version large files using Git Large File Storage (LFS;\textsuperscript{\protect\hyperlink{ref-11GtZ7icJ}{20}}/) text pointers, but may have to rely on external file storage alternatives (such as local or cloud-hosting).

@@

Laboratory notebooks help researchers track their research notes, methods, policies and protocols\textsuperscript{\protect\hyperlink{ref-wwHxTOtm}{21}}.
Virtual laboratory notebooks can be stored in GitHub repositories and provide the benefits of simultaneous, centralized and selective access, and allows for the easy update of policies and experiment protocols\textsuperscript{\protect\hyperlink{ref-10V7x4H4l}{22}}.
Researchers have been increasingly using GitHub to maintain versions and share digital laboratory notebooks\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3}} (\emph{e.g.}, \url{https://scheuerell-lab.github.io/lab-book}; \url{https://github.com/HuckleyLab/how_we_work}).

@@

Modern research in ecology and evolution is highly collaborative, bringing together multidisciplinary teams from various institutions.
On GitHub, collaborators can share feedback, brainstorm ideas, and troubleshoot problems (Figure \ref{fig:github-diagram}).
Project management can happen via three GitHub repository features: ``Issues'', ``Discussion'' and ``Projects''.
Github Issues allow for discrete tasks and sub-tasks to be identified, assigned to team members, and categorized with custom labels.
Github Discussions serve as a message board for conversation.
Finally, GitHub Projects integrate issues and pull requests on automated spreadsheets and project boards, providing users with real-time tracking of project priorities and status\textsuperscript{\protect\hyperlink{ref-RhBKe0MG}{23}}.
Scripts, commit messages, and pull requests can be linked directly to issues, discussions, and projects providing a clear record of project workflow.
The use of GitHub for all project-related conversation and planning, rather than e-mail or messaging tools, makes it easier to keep track of progress throughout the lifespan of a project.
Unlike emails and messages which can get lost as more new tasks arise, GitHub issues are intentionally closed by repository administrators hiding the issue from view (closed issues remain accessible but not immediately visible).
Project management in GitHub can also be integrated with third-party applications, such as ZenHub (https://www.zenhub.com/) or Slack (https://slack.github.com/).
ZenHub allows for the enhanced visualization and organization of repositories and their issues, while the GitHub for Slack integration allows notifications from GitHub events to be sent directly to users or group channels\textsuperscript{\protect\hyperlink{ref-1GRWGlDWy}{24}}.

@@

Making presentations, syllabi and other course materials can be done through most major high-level programming languages such as R, with \texttt{RMarkdown}\textsuperscript{\protect\hyperlink{ref-QqMezOMg}{25}}, Python, with \texttt{python-ppt} (\url{https://python-pptx.readthedocs.io}), and Julia, with \texttt{Remark.jl} (\url{https://juliapackages.com/p/remark}), and be version-controlled and stored in GitHub.
Once content is made, hosting a course website can be done through GitHub Pages\textsuperscript{\protect\hyperlink{ref-MDQfMwCW}{26}} (\emph{e.g.}, https://github.com/topics/course-website\textgreater).
This way, the course content can be available to enrolled students, as well as a global pool of learners and teachers interested in the course material.
Content can then be delivered via the course website, and/or a GitHub Organization with, for example, template repositories for assignments.
Student submissions are perhaps the most challenging component, but the new GitHub Classroom tool (\url{https://classroom.github.com}) allows instructors to host private assignments to be submitted collaboratively or individually as code or PDF files, and even build autograding tests.
Although time-consuming to establish, using these features can integrate learning version control and GitHub with the learning course content, and thus boost students feelings of self-efficacy and confidence\textsuperscript{\protect\hyperlink{ref-dqrFjoSb}{27}}.

@@

Personal or laboratory websites can improve the sharing of research findings, build online presence, and increase coordination of research efforts\textsuperscript{\protect\hyperlink{ref-HiIPSSHV}{28}}.
Despite many researchers in ecology and evolution having little experience in building or hosting webpages, many tools have been developed to help this process.
Static websites can now be easily built using independent software and languages, such as Jekyll (\url{https://jekyllrb.com}), Hugo (\url{https://gohugo.io}), Quarto (\url{https://quarto.org}), and wowchemy (\url{https://wowchemy.com}), or with the help of dependencies in the programming languages scientists commonly use, such as the \texttt{distill} (\url{https://github.com/rstudio/distill}) and \texttt{blogdown} (\url{https://bookdown.org/yihui/blogdown}) R packages.
The resulting files from static websites can then be hosted in repositories, from which one can activate GitHub's Pages (\url{https://pages.github.com}) feature, allowing for the direct live hosting of HyperText Markup Language (HTML), Cascading Style Sheets (CSS) and JavaScript files within a free github.io domain.
Website templates are readily available on GitHub (https://github.com/topics/website-template) or in user-specific repositories, which can be forked (\protect\hyperlink{definitions}{Box 1}) and customized\textsuperscript{\protect\hyperlink{ref-MXxgZJ45}{29}}.
Aside from free hosting services, GitHub Pages also allows websites to be autogenerated and instantly updated following content modifications\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}.
Creating and hosting websites on GitHub pages can be often time-consuming, as this process usually requires more knowledge in web design than out-of-the-box platforms (\emph{e.g.}, Wix, Weebly, Google Sites).
However, free hosting, widely available template customization, and versioning are strong advantages over the alternatives.

@@

Government, funding agencies, and publishers exercise rigorous open access data policies and mandates\textsuperscript{\protect\hyperlink{ref-PLmDFZrm}{30},\protect\hyperlink{ref-1Hcf13Q0k}{31}}.
However, code and data sharing may be met by individual reluctance, temporary embargoes, or partially prevented by privacy and confidentiality reasons\textsuperscript{\protect\hyperlink{ref-1Ch6LSHef}{32}--\protect\hyperlink{ref-1CzUZwyU2}{34}}.
Still, data deposition and ensuring its availability can amplify the outreach of published studies\textsuperscript{\protect\hyperlink{ref-666HppfO}{35}}, increase citation rates\textsuperscript{\protect\hyperlink{ref-1CcAUn3Lu}{36}}, and among many other reasons, enables the reproducibility and robustness of scientific advances\textsuperscript{\protect\hyperlink{ref-4LaijDIZ}{37}--\protect\hyperlink{ref-1HZdsK5Kn}{39}}.
While public repositories on GitHub make it easy to store and share data files, they are not considered long-term repositories for research materials.
This is because GitHub, a for-profit company, does not have long-term data availability guarantees, allowing users to delete or make repositories private after publication.
Also, GitHub does not issue Digital Object Identifiers (DOI) for content uploaded to their servers.
DOIs are persistent and citable unique alpha-numeric identifiers assigned to digitally stored research materials.
Because of this, scientists sharing code and data through GitHub are strongly encouraged to independently submit their research materials to long-term data archives (\emph{e.g.}, Zenodo, Figshare, Dryad, OSF\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-1Du6fzB8g}{40},\protect\hyperlink{ref-tbl:compare}{\textbf{tbl:compare?}}}.
Some of these options (Zenodo, Figshare and OSF) integrate with GitHub, allowing project, code, and data releases (\protect\hyperlink{definitions}{Box 1}) to be archived with versioned, citable DOIs.
Linking GitHub repositories with a DOI helps research become findable, properly cited, and can ensure long-term stability\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{41}}.

An important aspect of making code and data citable and reusable is to add an appropriate licence to protect intellectual property.
Code published without a licence is under exclusive copyright (by default), protecting it from copy, distribution, and modifications.
One may grant specific rights to their code for reuse by adding licensing files and specifications within GitHub repositories\textsuperscript{\protect\hyperlink{ref-TOsASkn5}{42}}.
The Choose a License (\url{https://choosealicense.com/non-software/}) website offers further guidance on the licenses available for research and creative products.
For example, Creative Commons (CC; \url{https://creativecommons.org/licenses/}) licenses can specify that shared code is intended for a specific analysis.
A CC BY 4.0 license specifies that any code (or other creative products) must be appropriately credited to its original author when distributed, adapted or reused.
A GNU General Public License v3 provides the same attributions above, but requires that all modified or adapted versions to be made public.

@@

Researchers can asynchronously communicate and individually or collaboratively work in GitHub.
By forking, branching, and cloning repositories (\protect\hyperlink{definitions}{Box 1}), researchers can simultaneously work on different issues of the same research project to then later merge changes to the main branch project with pull requests.
Explicit project organization and increased communication in GitHub Issues, Discussions, or within pull requests can help with project development and with potential merge conflicts due to users simultaneously working on the same sections\textsuperscript{\protect\hyperlink{ref-1BJcvyTmV}{43}}.
Moreover, version control\textsuperscript{\protect\hyperlink{ref-K7nbP1Ty}{44}} and commit reversal features allow researchers to track and progress without worrying about irreparably modifying someone else's work.
By enabling more comprehensive remote collaboration, GitHub encourages the exchange of ideas among researchers at different institutions and in different countries, which can serve to improve the quality of the research itself by providing open access to data and code.
In academic settings, GitHub can also facilitate interactions between research advisors and advisees, providing a platform for students or other trainees to share in-progress code, and flag specific challenges or questions for their supervisors or mentors (\protect\hyperlink{tbl:roles}{Table 1}).
Periodic code review\textsuperscript{\protect\hyperlink{ref-hm9PaCLD}{45}} can also help advisors identify errors early in the process, and inform further training and mentorship to fill gaps in skills.

@@

Conventional research practices typically rely on one or two people running and checking the data analyses, while most coauthors (and readers of the subsequent publication) see only the final results and a verbal description of the analytical steps.
In the developmental stages, collaborators can directly see the code for the analysis, manipulate and explore the data themselves, and check for errors.
Cynically, there is also more insurance against nefarious colleagues that may be tempted to distort results\textsuperscript{\protect\hyperlink{ref-18PTmKJkq}{49}}.
Collaborators as well as pre-publication and post-publication reviewers are better positioned to discover questionable findings if they have full and transparent access to the project.

>>>>

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/b78a03a54df81c17eb98b9ca80e842dcb1883ab1/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/b78a03a54df81c17eb98b9ca80e842dcb1883ab1}{SORTEE-Github-Hackathon/manuscript@b78a03a}
on June 24, 2022.

@@

Most scientists, including ecologists and evolutionary biologists, are increasingly dependent on computational tools in their research\textsuperscript{\protect\hyperlink{ref-fJWFe93e}{1}}.
Researchers write and use software packages or data analysis code (hereafter, code) to perform scientific tasks ranging from data management, data analysis, and study replication, to the application and the development of tools for hypothesis testing.
Maintaining code for scientific collaboration requires an efficient and well-documented work-flow\textsuperscript{\protect\hyperlink{ref-1Kqna6l2}{2}}.
To facilitate this process, scientists have been increasingly adopting tools from information and systems technology, such as cloud-based services for documentation and version control (\emph{e.g.}, from the Google Suite, the Microsoft Suite, and GitHub,\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3}}).
However, most researchers lack exposure to adequate code practices and thus dedicate valuable time and effort to self-teaching research-facilitating tools.
Thus, researchers may not adhere to standards of code quality and maintenance\textsuperscript{\protect\hyperlink{ref-fJWFe93e}{1},\protect\hyperlink{ref-10SpoByIw}{4},\protect\hyperlink{ref-O6UbstGG}{5}}.
Here, we review and discuss one of the most used web-based platforms for computational version control and collaboration, GitHub, and provide researchers in ecology and evolutionary biology (EEB) with practical workflows to facilitate and improve their code and its management.

With over 73 million registered users as of 2022, GitHub is the most widely-used web platform for collaborating on computer code\textsuperscript{\protect\hyperlink{ref-nwCtHDCn}{6}}.
GitHub provides a simplified but powerful web interface that allows users to participate in projects, contribute code, report and discuss code bugs, discover existing code and data, and publish new code.
Through version control, users have a detailed, chronological record of the files and directories stored in their repositories (see \protect\hyperlink{definitions}{Box 1};\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}).
This workflow provides a strong and clear advantage over sending files back-and-forth (\emph{e.g.} via email), a process that can become challenging and time-consuming in more long-term and collaborative projects\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8}}.
Through its combination of version control and and collaborative features, GitHub facilitates open source code alongside collaborative development\textsuperscript{\protect\hyperlink{ref-kEX5dgzK}{9}}.

Git is the version control system that enables all the collaborative tools available on GitHub.
Although the understanding of basic concepts of Git (such as commit, push, pull, checkout; see \protect\hyperlink{definitions}{Box 1}) is necessary, the GitHub web-based platform and its integrated development environments (such as the GitHub Desktop) allow users to perform most repository and data management operations without opening Git command-line sessions.

The expansive GitHub user-community and numerous GitHub resources have boosted its popularity\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-13jOlVcpp}{10}/,\protect\hyperlink{ref-u5aEVE4B}{11}}.
Nevertheless, although multiple articles have encouraged researchers in EEB to adopt GitHub as part of their research process\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-3DKwn1sY}{12}}, its use is still not widespread.
First-time users without formal training in information technology may face steep learning curves because GitHub and its features have been centered on collaboration for code development in information systems\textsuperscript{\protect\hyperlink{ref-139b0pSGc}{13}}.
Moreover, domain-specific resources providing tractable examples and practical guidance for researchers in EEB on GitHub are scarce (but see \url{https://ourcodingclub.github.io}; \url{https://www.openscapes.org}).
Widespread adoption of GitHub for collaborating on research tasks can ultimately enable EEB researchers to spend less time on creating novel processes for collaboration and more time on their research\textsuperscript{\protect\hyperlink{ref-ydrk01SR}{14}}.
More importantly, increasing the availability of data and code management standards -- of which GitHub is one increasingly important component -- makes research more reproducible and collaborative\textsuperscript{\protect\hyperlink{ref-13QX8XU3J}{15}}.

This paper is the result of an academic hackathon held during the 2021 conference for the Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology (SORTEE, \url{https://www.sortee.org}).
We convened a group of around thirty researchers in EEB with varying levels of familiarity with using GitHub as part of their research projects to showcase and discuss how existing features can contribute to documentation and collaboration in EEB research.
During the hackathon, we identified the need for a formal discussion on how EEB researchers can benefit from GitHub and its features to make research more collaborative and transparent.
Here, we outline twelve practical ways that EEB researchers can use GitHub features for more collaborative, transparent, and reproducible science.
We also provide critical perspectives on features that could be improved and catered towards research development.

@@

\textbf{fork}: A fork is a copy of a \textbf{repository} hosted on GitHub. If a repository is public, then anyone can make a fork. Even if they do not have access to \textbf{push} to the original repository, they can make a fork and edit it independently. Forks are linked to the original GitHub repository and ``upstream'' changes (\emph{i.e.} those in the original repository) can be \textbf{merged} to keep the fork up to date with the original project. Changes made in the fork can be integrated into the original project via \textbf{pull requests}. \\
\textbf{clone}: Cloning a \textbf{repository} is a way of making a local copy (\emph{i.e.} on your computer) of a GitHub \textbf{repository}. If you have access to \textbf{push} to a \textbf{repository}, this can be a first step to contributing to a project. \\

@@

\textbf{commit}: Commits are snapshots of the development of a project. In Git, versions of files and directories are uniquely identified as ``commits'', allowing one to identify and track modifications line-by-line. Commits can include changes in multiple files and must include a brief commit message describing the changes made. A typical workflow is to make some related changes in files, add a commit message (\emph{e.g.} ``generate and include fig1 in results''), and after several commits \textbf{push} those commits to the remote (\emph{i.e.}, cloud-based) GitHub \textbf{repository}. \\

@@

An EEB research compendium includes all computational materials related to research production, including data, code for analyses and protocols.
Having copies of these files safely stored is essential to protect against accidental modifications or deletions.
Many researchers begin using GitHub as a means to store (or backup) their research compendium\textsuperscript{\protect\hyperlink{ref-MwwMapRG}{16}} to a centralized, readily-available remote server (see \protect\hyperlink{definitions}{Box 1}).
A centralized research compendium stored in a version-controlled repository has the advantages of facilitating collaboration, integrating data and code archiving services, allowing file versions to be accessed and restored, and contributes to open science (see sections below).

GitHub limits committed file sizes to 100 Mb (megabytes) (\textsuperscript{\protect\hyperlink{ref-1Co6ZZjF1}{17}}), which can make it challenging for centralizing research compedia containing larger file sizes.
Users may still version large files using Git Large File Storage (LFS;\textsuperscript{\protect\hyperlink{ref-11GtZ7icJ}{18}}/) text pointers, but may have to rely on external file storage alternatives (such as local or cloud-hosting).

@@

Laboratory notebooks help researchers track their research notes, methods, policies and protocols\textsuperscript{\protect\hyperlink{ref-wwHxTOtm}{19}}.
Virtual laboratory notebooks can be stored in GitHub repositories and provide the benefits of simultaneous, centralized and selective access, and allows for the easy update of policies and experiment protocols\textsuperscript{\protect\hyperlink{ref-10V7x4H4l}{20}}.
Researchers have been increasingly using GitHub to maintain versions and share digital laboratory notebooks\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3}} (\emph{e.g.}, \url{https://scheuerell-lab.github.io/lab-book}; \url{https://github.com/HuckleyLab/how_we_work}).

@@

Modern research in ecology and evolution is highly collaborative, bringing together multidisciplinary teams from various institutions.
On GitHub, collaborators can share feedback, brainstorm ideas, and troubleshoot problems (Figure \ref{fig:github-diagram}).
Project management can happen via three GitHub repository features: ``Issues'', ``Discussion'' and ``Projects''.
Github Issues allow for discrete tasks and sub-tasks to be identified, assigned to team members, and categorized with custom labels.
Github Discussions serve as a message board for conversation.
Finally, GitHub Projects integrate issues and pull requests on automated spreadsheets and project boards, providing users with real-time tracking of project priorities and status\textsuperscript{\protect\hyperlink{ref-RhBKe0MG}{21}}.
Scripts, commit messages, and pull requests can be linked directly to issues, discussions, and projects providing a clear record of project workflow.
The use of GitHub for all project-related conversation and planning, rather than e-mail or messaging tools, makes it easier to keep track of progress throughout the lifespan of a project.
Unlike emails and messages which can get lost as more new tasks arise, GitHub issues are intentionally closed by repository administrators hiding the issue from view (closed issues remain accessible but not immediately visible).
Project management in GitHub can also be integrated with third-party applications, such as ZenHub (https://www.zenhub.com/) or Slack (https://slack.github.com/).
ZenHub allows for the enhanced visualization and organization of repositories and their issues, while the GitHub for Slack integration allows notifications from GitHub events to be sent directly to users or group channels\textsuperscript{\protect\hyperlink{ref-1GRWGlDWy}{22}}.

@@

Making presentations, syllabi and other course materials can be done through most major high-level programming languages such as R, with \texttt{RMarkdown}\textsuperscript{\protect\hyperlink{ref-QqMezOMg}{23}}, Python, with \texttt{python-ppt} (\url{https://python-pptx.readthedocs.io}), and Julia, with \texttt{Remark.jl} (\url{https://juliapackages.com/p/remark}), and be version-controlled and stored in GitHub.
Once content is made, hosting a course website can be done through GitHub Pages\textsuperscript{\protect\hyperlink{ref-MDQfMwCW}{24}} (\emph{e.g.}, https://github.com/topics/course-website\textgreater).
This way, the course content can be available to enrolled students, as well as a global pool of learners and teachers interested in the course material.
Content can then be delivered via the course website, and/or a GitHub Organization with, for example, template repositories for assignments.
Student submissions are perhaps the most challenging component, but the new GitHub Classroom tool (\url{https://classroom.github.com}) allows instructors to host private assignments to be submitted collaboratively or individually as code or PDF files, and even build autograding tests.
Although time-consuming to establish, using these features can integrate learning version control and GitHub with the learning course content, and thus boost students feelings of self-efficacy and confidence\textsuperscript{\protect\hyperlink{ref-dqrFjoSb}{25}}.

@@

Personal or laboratory websites can improve the sharing of research findings, build online presence, and increase coordination of research efforts\textsuperscript{\protect\hyperlink{ref-HiIPSSHV}{26}}.
Despite many researchers in ecology and evolution having little experience in building or hosting webpages, many tools have been developed to help this process.
Static websites can now be easily built using independent software and languages, such as Jekyll (\url{https://jekyllrb.com}), Hugo (\url{https://gohugo.io}), Quarto (\url{https://quarto.org}), and wowchemy (\url{https://wowchemy.com}), or with the help of dependencies in the programming languages scientists commonly use, such as the \texttt{distill} (\url{https://github.com/rstudio/distill}) and \texttt{blogdown} (\url{https://bookdown.org/yihui/blogdown}) R packages.
The resulting files from static websites can then be hosted in repositories, from which one can activate GitHub's Pages (\url{https://pages.github.com}) feature, allowing for the direct live hosting of HyperText Markup Language (HTML), Cascading Style Sheets (CSS) and JavaScript files within a free github.io domain.
Website templates are readily available on GitHub (https://github.com/topics/website-template) or in user-specific repositories, which can be forked (\protect\hyperlink{definitions}{Box 1}) and customized\textsuperscript{\protect\hyperlink{ref-MXxgZJ45}{27}}.
Aside from free hosting services, GitHub Pages also allows websites to be autogenerated and instantly updated following content modifications\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}.
Creating and hosting websites on GitHub pages can be often time-consuming, as this process usually requires more knowledge in web design than out-of-the-box platforms (\emph{e.g.}, Wix, Weebly, Google Sites).
However, free hosting, widely available template customization, and versioning are strong advantages over the alternatives.

@@

Government, funding agencies, and publishers exercise rigorous open access data policies and mandates\textsuperscript{\protect\hyperlink{ref-PLmDFZrm}{28},\protect\hyperlink{ref-1Hcf13Q0k}{29}}.
However, code and data sharing may be met by individual reluctance, temporary embargoes, or partially prevented by privacy and confidentiality reasons\textsuperscript{\protect\hyperlink{ref-1Ch6LSHef}{30}--\protect\hyperlink{ref-1CzUZwyU2}{32}}.
Still, data deposition and ensuring its availability can amplify the outreach of published studies\textsuperscript{\protect\hyperlink{ref-666HppfO}{33}}, increase citation rates\textsuperscript{\protect\hyperlink{ref-1CcAUn3Lu}{34}}, and among many other reasons, enables the reproducibility and robustness of scientific advances\textsuperscript{\protect\hyperlink{ref-4LaijDIZ}{35}--\protect\hyperlink{ref-1HZdsK5Kn}{37}}.
While public repositories on GitHub make it easy to store and share data files, they are not considered long-term repositories for research materials.
This is because GitHub, a for-profit company, does not have long-term data availability guarantees, allowing users to delete or make repositories private after publication.
Also, GitHub does not issue Digital Object Identifiers (DOI) for content uploaded to their servers.
DOIs are persistent and citable unique alpha-numeric identifiers assigned to digitally stored research materials.
Because of this, scientists sharing code and data through GitHub are strongly encouraged to independently submit their research materials to long-term data archives (\emph{e.g.}, Zenodo, Figshare, Dryad, OSF\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-1Du6fzB8g}{38},\protect\hyperlink{ref-tbl:compare}{\textbf{tbl:compare?}}}.
Some of these options (Zenodo, Figshare and OSF) integrate with GitHub, allowing project, code, and data releases (\protect\hyperlink{definitions}{Box 1}) to be archived with versioned, citable DOIs.
Linking GitHub repositories with a DOI helps research become findable, properly cited, and can ensure long-term stability\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{39}}.

An important aspect of making code and data citable and reusable is to add an appropriate licence to protect intellectual property.
Code published without a licence is under exclusive copyright (by default), protecting it from copy, distribution, and modifications.
One may grant specific rights to their code for reuse by adding licensing files and specifications within GitHub repositories\textsuperscript{\protect\hyperlink{ref-TOsASkn5}{40}}.
The Choose a License (\url{https://choosealicense.com/non-software/}) website offers further guidance on the licenses available for research and creative products.
For example, Creative Commons (CC; \url{https://creativecommons.org/licenses/}) licenses can specify that shared code is intended for a specific analysis.
A CC BY 4.0 license specifies that any code (or other creative products) must be appropriately credited to its original author when distributed, adapted or reused.
A GNU General Public License v3 provides the same attributions above, but requires that all modified or adapted versions to be made public.

@@

Researchers can asynchronously communicate and individually or collaboratively work in GitHub.
By forking, branching, and cloning repositories (\protect\hyperlink{definitions}{Box 1}), researchers can simultaneously work on different issues of the same research project to then later merge changes to the main branch project with pull requests.
Explicit project organization and increased communication in GitHub Issues, Discussions, or within pull requests can help with project development and with potential merge conflicts due to users simultaneously working on the same sections\textsuperscript{\protect\hyperlink{ref-1BJcvyTmV}{41}}.
Moreover, version control\textsuperscript{\protect\hyperlink{ref-K7nbP1Ty}{42}} and commit reversal features allow researchers to track and progress without worrying about irreparably modifying someone else's work.
By enabling more comprehensive remote collaboration, GitHub encourages the exchange of ideas among researchers at different institutions and in different countries, which can serve to improve the quality of the research itself by providing open access to data and code.
In academic settings, GitHub can also facilitate interactions between research advisors and advisees, providing a platform for students or other trainees to share in-progress code, and flag specific challenges or questions for their supervisors or mentors (\protect\hyperlink{tbl:roles}{Table 1}).
Periodic code review\textsuperscript{\protect\hyperlink{ref-hm9PaCLD}{43}} can also help advisors identify errors early in the process, and inform further training and mentorship to fill gaps in skills.

@@

Conventional research practices typically rely on one or two people running and checking the data analyses, while most coauthors (and readers of the subsequent publication) see only the final results and a verbal description of the analytical steps.
In the developmental stages, collaborators can directly see the code for the analysis, manipulate and explore the data themselves, and check for errors.
Cynically, there is also more insurance against nefarious colleagues that may be tempted to distort results\textsuperscript{\protect\hyperlink{ref-18PTmKJkq}{47}}.
Collaborators as well as pre-publication and post-publication reviewers are better positioned to discover questionable findings if they have full and transparent access to the project.
